---
title: "Chi_squared"
author: "Tara Doma Lama"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#load the libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggstatsplot)

```

#STEP 1: 

load the file
give new column names
tidy the data
```{r}
conopeptide <- read_csv("~/CSB-BIOL425/unix/sandbox/conopeptide_csv_final", col_names = FALSE)
colnames(conopeptide) <- c("id", "hmm", "psm", "predicted")#new column names
conopeptide_new <- conopeptide %>% 
  mutate(seqid = str_split_i(id, "_", 1), framework = str_split_i(id, "_", 2))  
  glimpse(conopeptide_new)
```

#Step 2
Visaulize the data
```{r}
fam_counts <- conopeptide_new %>% 
  group_by(predicted, framework) %>% 
  count()
head(fam_counts)

#plot the data
fam_counts %>% 
  ggplot(aes(x = framework, y = n))+
  geom_col()+
  theme_bw()+
  facet_wrap(~predicted)
```

#Visualize separately
```{r}
conopeptide_family <- conopeptide_new %>%
  filter(predicted != c("UNKNOWN")) %>% 
  group_by(predicted) %>%
  count()
# literature says M is the most observed superfamily and here it comes second to the O1
conopeptide_family %>% 
  ggplot(aes(x = predicted, y = n))+
  geom_col()+
  theme_bw()+
  coord_flip()
# literature says cysteine framework is the most observed framework.

```


```{r}
conopeptide_framework <- conopeptide_new %>%
  group_by(framework) %>%
  count() %>% 
  arrange(desc(n))
  

conopeptide_framework %>% 
  ggplot(aes(x = framework, y = n))+
  geom_col()+#default stat is identity, height of the bar represents the raw value of the y
  theme_bw()
# literature says cysteine framework 6, 1  is the most observed framework.

```

#most enocded superfamilies is A, M and O1
```{r}
A <- function(x = "Novel"){
  conopeptide_new %>%
    filter(framework == x) %>% 
    group_by(predicted, framework) %>%
    count()}

A( x = "Novel")
A( x = "M")#M contains nine cysteine frameworks and they are found in all other superfamilies, look at framework 14 and 9
A( x = "O1")

```

take pedicted and framework as the two variables
visaulize their association via the plot
```{r}
#get the counts
fam_counts_A <- fam_counts %>% # new object 
  filter(predicted %in% c("A", "M")) %>% 
  filter(framework %in% c(8,22)) %>% #framework 8 is the pharmacological interest
  group_by(predicted, framework) %>% 
  summarise(sum_total = sum(n))
fam_counts_A

#plot the graph
fam_counts_A %>% 
  ggplot(aes(x = framework, y = sum_total, fill = predicted))+
  geom_col(stat = "identity", position = "dodge")
  

```
#STEP : 3

chissquare is non parametric, calculate the frequencies of the pairwise association
conditions for chisquared test are:
1.two categorical variables: framework, predicted
2.there are two categories for each group
3. frequencies in each cell is at least one
4. independence of the sample
Since, we meet all the above condition, we perform the chi squared test

Null hypothesis: there is no association between the predicted, superfamily, and framework.

Chi_square test


```{r}
fam_count_clean <- xtabs (sum_total ~ predicted + framework, data = fam_counts_A)# for correct format
fam_count_clean

#chisquare test
fam_count_chi <- chisq.test(fam_count_clean)

fam_count_chi
fam_count_chi$expected
fam_count_chi$residuals# see where the observation lies compared to the expected


```
#Here, the p value indicates that the probability of observing the chisquared test statistics at least as above or larger is less than 2.2e-16. Hence, we reject the null hpothesis.


#STEP 4:
#replot with the statistical result
```{r}
fam_counts <- conopeptide_new %>%
  group_by(predicted, framework) %>% #grouped data should be ungrouped
  filter(predicted %in% c("A", "M")) %>% 
  filter(framework %in% c(8,22))
# if the frequency distribution of two categorical variables are independent of each other using the contingency table analysis
fam_stats <- ggbarstats(
  data = fam_counts,
  x = predicted ,# rows in contingency table are mapped in x axis
  y = framework,
  type = "nonparametric",
  label = "percentage" # both gives counts and percentage
)
fam_stats

#get the stats
extract_stats(fam_stats)
```
# because the type of the test is non parametric, they performed the pearson's Chi_squared test.
#yate's continuity decreases chisquared value.







